---
title: "Introduction to rbst"
author: "Tarak Shah"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rbst}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
### Background

`rbst` is a purely functional implementation of a balanced search tree. In addition to the normal advantages of a balanced search tree (guaranteed fast retrieval, deletion, and insertion of new key-value pairs, ordered search operations), `rbst` also has the (natural in `R`) quality of non-destructive updates. 

```{r examples}
library(rbst)
example <- bst(keys = letters, values = seq_along(letters))
example

# retrieval:
example["t"]

# ordered search for keys:
keys_between(example, "f", "i")

# non-destructive insertions:
modified_example <- insert(example, "new key", 97)
contains(modified_example, "new key")
contains(example, "new key")
```

`delete` functions also allow for priority queue-like behavior:

```{r delmin}
example <- delete_min(example)
contains(example, "a")
example <- delete_min(example)
contains(example, "a")
contains(example, "b")
```

-----

### References

Okasaki, Chris. *Purely Functional Data Structures.* Cambridge University Press, 1999.

O'Neil, Shawn T. "Implementing Persistent O(1) Stacks and Queues in R." *The R Journal* Volume 7/1 (June 2015).

Oster, Julien. "An Agda implementation of deletion in Left-leaning Red-Black trees." http://www.reinference.net/llrb-delete-julien-oster.pdf

Sedgewick, Robert and Kevin Wayne. *Algorithms, 4th Edition.* Addison Wesley Professional, 2011.